### 2024-08-30 ms.qmd
Updates

---
title: "Early environmental conditions do not impact behavioural flexibility in an invasive and non-invasive lizard species"
format:
  docx: 
    bibliography: "./bib/refs.bib"
    csl: "./bib/animal-behaviour.csl"
    reference-doc: "./bib/tmpl.docx"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
```

Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Kristoffer H. Wild^1,3^, Chris Friesen^4^, Basile Mauclaire^1,5^, Amelia Y. Peardon ^1^, Daniel W.A. Noble^1^

^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
^3^ Melbourne University, Melbourne, VIC 3000, Australia  
^4^ University of Wollongong, Wollongong, NSW 2500, Australia  
^5^ Université de Lille, Lille 59000, France  
$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

## Abstract
Behavioural flexibility is expected to be crucial for animals adapting to environmental changes. However, the conditions experienced during early development can profoundly impact behavioural flexibility making it unclear how populations will respond to novel circumstances. Stressful situations faced by the parents can have a direct impact on offspring cognition through the transmission of glucocorticoids - stress-related hormones that affect offspring cognition. At the same time, stressful conditions can influence parental behaviour during nesting and consequently the thermal developmental conditions offspring experience. Here, we investigated the interactive effects of prenatal corticosterone (CORT) and temperature on behavioural flexibility in two lizard species, _Lampropholis delicata_ and _L. guichenoti_. We manipulated prenatal CORT levels and incubation temperature in a 2x2 factorial design, and then assessed behavioural flexibility through a reversal learning task. We hypothesized prenatal CORT and cold temperatures to impair with lizards' performance in the reversal task. Given _L. delicata_'s success as an invassive species, we expected this species to behave more flexibly and be less affected by early environmental conditions. Contrary to our expectations, behavioural flexibility appears to be robust to prenatal temperature and CORT in both species. The lack of difference in reversal learning between _L. delicata_ and _L. guichenoti_ suggests that novel environments are unlikely to influence flexible behavioural learning and that behavioural flexibility itself is unlikely to explain differences in invasion success between these species.

## Introduction
Behavioural flexibility describes the ability of individuals to adjust behaviour adaptively in response to changes in the internal or external environment [@brown2010behavioral]. It comprises a suite of behaviours, including problem-solving or innovation [@brown2010behavioral], that can be beneficial for adapting to novel circumstances [@szabo2020can; @szulkin2020urban]. Species with greater behavioural flexibility are expected to be more resilient to environmental changes because behaving flexibly is predicted to be crucial for exploiting new resources and facing new challenges [@wright2010behavioral; @chapple2012can; @szabo2020can; @szulkin2020urban]. For example, successful colonisation of new environments by birds is related to higher frequency of foraging innovations [@sol2000behavioural]. Although there is limited understanding of the physiological mechanisms involved in animals' ability to behave flexibly, relative brain size has been shown to be a good predictor of behavioural flexibility in mammals, birds, reptiles and amphibians [@sol2000behavioural; @sol2008brain; @amiel_smart_2011]. This suggests that behavioural flexibility is likely to be underpinned by complex neural mechanisms, and that alterations in these mechanisms may impact animals' ability to respond to novel situations. The brain is considered especially sensitive to environmental changes during the early stages of development [@zhu_prenatal_2004], creating uncertainty about how the conditions faced during early life might impact behavioural flexibility.

The conditions in which organisms develop can be altered by the circumstances faced by their parents. For instance, exposing three-spined stickleback (_Gasterosteus aculeatus_) females to a model predator during ovulation, affected their offspring's tedency to use social cues [@feng2015effect]. Under stressful situations, animal's sensory systems detect and transmit environmental signals to instigate adaptive physiological and behavioural adjustments – responses that are mediated by glucocorticoids (GCs) [@sapolsky_how_2000]. Glucocorticoids can be transmitted to offspring and influence their phenotype and development through transgenerational effects [@Crino_2023]. In addition to the environments experienced by parents, offspring also experience potentially stressful environmental conditions that can interact with or amplify parental effects. For example, human disturbance alters snapping turtles' nest-site choice, ultimately impacting the temperature at which eggs are incubated [@kolbe2002impact]. Animals’ responses to abrupt temperature changes are mediated by GCs [@Crino_2023], which can influence animals’ decision-making while nesting [@kolbe2002impact]. As a result, GCs transmission and early thermal environment are expected to interact and shape offspring traits. Both GCs and temperature are known to exert potent effects on brain development [@jonson1976effect; @coomber_independent_1997; @zhu_prenatal_2004; @amiel_effects_2017] and cognition [@clark_colour_2014; @lui2017chronic], but little is known about how they may interact to affect behavioural flexibility. Since behaving flexibly is essential for facing novel environments, studying the interactive effect of early environment on behavioural flexibility is crucial if we aim to understand animals' capability to adapt to environmental changes.

Here, we studied the effects of prenatal temperature and corticosterone (CORT) (the main GC in reptiles) [@Crino_2023], on behavioural flexibility in two species of reptiles, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). Both species have similar life-histories and overlapping distributions [@chapple_know_2011; @chapple_biology_2014], but only _Lampropholis delicata_ has been successful in colonising different areas around the globe [@baker1979rainbow; @chapple2013divergent; @chapple2015deliinvLHI]. Previous studies exploring behavioral differences between the two species have found _L. delicata_ to be more exploratory than _L. guichenoti_ [@chapple_know_2011], but no differences in learning were observed between the skinks in an associative learning task [@bezzina2014does]. Disparities in behavioural flexibility may be driving the differences in invasion success between both skinks, a prediction supported in other invasive species [@sol2000behavioural; @sol2008brain; @wright2010behavioral; @amiel_smart_2011; @chapple2012can]. However, the relative impact of early environments on behavioural flexibility remains unexplored. Early environments are known to affect suites of morphological and behavioural traits differently in each species [@carrasco2023maternal; @kar2023heritability], but uncertainty still exists about the interactive role of maternal CORT and early temperatures on behavioural flexibility. We manipulated CORT and incubation temperature in both species and then subjected hatchlings to a reversal learning task. Reversal learning, in particular, is a widely employed tool to measure behavioral flexibility [@gapp2014early; @hurtubise2017effects], as it assesses an individual's ability to alter a learned behavior in response to environmental changes, providing researchers with a clear indicator of their ability to adjust to new conditions [@brown2010behavioral].

We predicted that both cold incubation temperatures and CORT treatment will impair behavioural flexibility [@clark_colour_2014; @lui2017chronic], and the combination of both treatments to have the most detrimental effect on behavioural flexibility. Since invassive species are expected to behave more flexibly [@sol2000behavioural; @sol2008brain; @wright2010behavioral; @amiel_smart_2011; @chapple2012can], we predicted that _L. delicata_, on average, will perform more proficiently in the task compared to _L. guichenoti_ and that behavioural flexibility in _L. delicata_ will be more robust to early environmental conditions.

## Methods

#### Husbandry  
*Breeding colony* -- We tested juveniles originating from a breeding colony established in the lab since 2019. A total of 270 and 180 adults of *L. delicata* and *L. guichenoti*, respectively, were established in containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure. Enclosures had non-stick matting, a shelter, and several small water dishes. Water was given daily, and they were fed approx. 40 mid-size crickets (*Acheta domestica*) per enclosure three days a week. Crickets were dusted with calcium weekly and multivitamin and calcium biweekly. Using a heat chord and a heat lamp we created a temperature gradient and kept lights on a 12 h light:12 h dark cycle. UVA/UVB lighting was also provided. Room temperatures were set to 22-24ºC, and the warm side of enclosures reached 32ºC.

*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside that were placed on one side of the communal enclosures (see above). We checked for the presence of eggs in the boxes three days a week. After collection, we measured the length and width of eggs with a digital caliper to the nearest 0.1 mm and weighed them with a digital scale ± 0.001g error (OHAUS, Model spx123). We also assigned identities to the clutch and each egg. Eggs were then treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.

*Hatchlings* – Eggs in the incubator were checked three times a week for hatchlings. After hatchling, we measured juveniles’ SVL and Tail Length (TL) with a ruler to the nearest mm and weighed them with a digital scale ± 0.001g error. We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting, a shelter, and a small water dish. During this period, they were sprayed with water every day and received 3-6 small _A. domestica_ crickets three times a week. All care otherwise followed similar protocols to adults (see above).
Two weeks before starting the training phase (see below), lizards were moved to an experimental arena for acclimatisation. The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the side and a water dish on the other. These new enclosures were placed in two rooms across 7 different racks that were monitored by 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below). Conditions in these rooms where identical to the main room, but the number of lizards per species and treatment in each rack was counterbalanced to control for any potential effect of the room or the position of the lizard on the rack. During acclimatisation and throughout the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied _ad libitum_. We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle. The temperature of both rooms was set to between 22-24ºC.

#### Manipulating Early Thermal and CORT Environments

To empirically test the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A). We first allocated eggs into one of two different treatments: in the CORT treatment, eggs were topically supplied with 5µL of CORT dissolved in 100% Ethanol (vehicle) at a final concentration of 10 pg CORT/mL; in the Control treatment eggs received an equal volume of the vehicle. CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data). Doses were determined by referencing published yolk corticosterone concentrations in other oviparous species [@lovern2008effects ; @hanover2019corticosterone] and our own unpublished data. After the hormone treatment, were then incubated in one of the two previously mentioned temperature regimes (Cold or Hot) until hatching. These temperatures represent the upper and lower limit of the natural incubation temperatures [@qualls2000post; @cheetham2011embryonic]. The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.

#### Quantifying Behavioural Flexibility  
The learning test was divided in three main phases: a habituation phase where lizards had to learn to eat a frozen-thawed cricket from white 3D-printed polylactic acid (PLA) ramps (9 L x 4 W x 5 H cm); a colour association task where lizards had to associate the food reward with one of two colours (red or blue) (@fig-Methods Associative task); and finally, to quantify behavioural flexibility, we reversed the previously trained colour (red or blue) so that the lizards needed to learn to associate the food with the opposite colour (@fig-Methods Reversal task).

The habituation phase was divided into three stages: in the first stage, the reward, a small, frozen cricket (*A. domestica*) was placed in an opaque petri dish (3D x 1.6H cm) (@fig-Methods Habituation phase, Stage 1); in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps (@fig-Methods Habituation phase, Stage 2); and in the third stage, the cricket was left inside a well (3D x 1.75H cm) on top of the ramp (Fig. @fig-Methods Habituation phase, Stage 3). Trials in every stage lasted one hour, begining when the feeding block (petri dish, ramp, or both) was placed inside the enclosure and finishing when we took the feeding block away. We recorded whether the cricket had been consumed or not and we considered a stage as completed if the lizards ate the crickets in at least 5 out of 6 trials. This phase lasted 38 days, and only in one case a lizard did not progress to the next phases because it did not pass the aqcuisition criteria.

In the colour association (Associative task in Fig. @fig-Methods) and reversal task (Reversal task in @fig-Methods B), we used three ramps that were identical in shape and size but different in the colour. The colours of these ramps were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. However, to control for potential colour biases, we assigned one group of lizards within each treatment to associate food in the blue ramp followed by the red ramp in the reversal, while in the other group the order was reversed. As in the last stage of the habituation phase, we placed the cricket inside the wells of the ramps, but then two of them were covered with 3D-printed lids (3D x 0.5H cm) so prey was only accessible on one ramp. The food reward was placed on all three ramps to avoid lizards using prey chemical cues that were released in the closed wells through a series of small holes on the top of the lids. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. Lizards were tested in the association task once a day for 35 days and once a day for 40 days in the reversal.

The full experiment was performed between the 6^th^ of March until the 26^th^ of June 2023. Tests were made between 11 to 12 am, when the lizards were active. Trials in the learning phases (colour associative task and reversal tasks) were recorded with different CCTV systems always using the same camera per individual. We recorded whether the animal chose the correct ramp first or not. We considered that a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording. Those trials were scored as NA. Here, we analysed the data of the reversal task as a measure of behavioural flexibility [@brown2010behavioral].

```{r, fig-Methods}
#| label: fig-Methods
#| fig.cap: "Experimental design of early environment manipulation and learning tasks. Panel A represents the early environment manipulation for both species. Panel B shows the habituation phase with the respective three different stages. And panel C represents the associative and reversal tasks; white lids show the ramps where the food reward was not accessible."

knitr::include_graphics("./Others/BEHFLEX_FIG_1.png")

```

#### Statistical analyses  
We performed the analyses for each species separately. We fitted two different Bayesian multilevel models using the brm function from the _brms_ package [@burkner2017brms] using an R extension (version 2.8.2) [@R]. We ran four parallel MCMC chains of 3000 iterations for each model, with a warmup period of 1000 iterations. We modelled correct choices [correct (1) or not (0)] as the response variable, and trial, hormone (CORT versus Control), incubation temperature (Cold versus Hot), and the three-way interaction between trial, hormone, and temperature. The error structure was modeled using a Bernoulli distribution with a logit link function (family = Bernoulli(link = 'logit')). We included a random intercept and slope (trial) for each lizard in our models. We also incorporated the clutch identity as a random factor. _L. delicata_ lays one clutch per year, while _L. guichenoti_ lays two [@chapple_know_2011; @chapple_biology_2014]. Since eggs were collected during half of the breeding season, clutches likely come from different mothers. By including clutch as a random factor, maternal condition should be accounted for in our analyses. Additionally, differential paternity may be captured by including both clutch and lizard identity in the model. Learning can also be age-dependent (see @noble_age-dependent_2014), and given that lizards incubated at different temperatures hatch at different times we explored the effects of age on learning. Indeed, lizards’ age at the beginning of the experiment ranged from 41 to 148 days old in _L. delicata_ and 48 to 132 in _L. guichenoti_. However, when this variable was included in the models we did not find any significant effect of age (see Supplementary material). As such, we present models without age as a fixed effect. We used the posterior distributions of parameters from these models to test for differences in learning rate between treatments and species. Learning slopes were obtained using the ‘trial’ estimates and its interaction with hormone and temperature treatments. Slope estimates greater than zero were considered as evidence of learning, while those less or equal to zero were not. We calculated the probability that slopes or contrasts between slopes differed from zero (p~mcmc~) using the posterior distributions. We considered statistical significance if p~mcmc~ < 0.05.

#### Ethical note
We collected subjects from a breeding colony kept in the lab since 2019. Both the breeding animals and the ones used in this experiment were provided humane laboratory housing, with thermoregulatory opportunities, light (UV and heat) and moderate levels of humidity (see above for details). Lizards in the colony received food ad libitum while the subjects in the experiment were feed one cricket daily during the experiment. All lizards were handled minimally to avoid stress. Experimental procedures and laboratory housing complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33). Animals born in captivity cannot be re-released into the wild as stringent biosafety protocols across the states prohibit this. It is also not possible to rehome such a large number of animals. Therefore, upon completion of experimental trials, lizards were kept in their enclosure until they were humanely euthanized following approved procedures. To euthanise animals, we use an injection of sodium pentobarbital (i.e., Lethabarb) (diluted 1:4 ringers solution) at 60-100mg/kg, injected intrapleuroperitoneally. We monitored the animals to ensure there was no irritation from the agent as indicated by distressed animals. Gently pinching the toes was used to confirm the absence of a response before the animal was disposed of.  
Crickets were maintained in big communal enclosures in the same room where the big colony is, and at the same temperature and light conditions. They were provided vegetables and water ad libitum, and they were given several egg carton shelters to refuge. All lizards were fed alive crickets except during the experiment, when we used crickets frozen at -21 ºC for 24 hours.

## Results

```{r, cleandata}
#| label: cleandata
# Obtain the main df using "./R/1_data_process.R"
source(here("R", "1_data_process.R"))
```

```{r, sampleSize}
#| label: sampleSize
# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
source(here("R", "func.R"))
#
specie <- c("delicata", "guichenoti")
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
#
n_list <- list()
#
for(i in 1:length(specie)){
  for(k in 1:length(hormone)){
    for(l in 1:length(temperature)){
      n_list[[paste(specie[i], hormone[k], temperature[l], sep = "_")]] <- sample(clean_df, specie[i], hormone[k], temperature[l])
      }
    }
  }
```

We started with 96 lizards, 48 per species and 12 per treatment per species. However, our final sample size was 81 due to natural mortality (n = 11), failure to complete the training stage (n = 1), or because the number of failed tasks (see above) was higher than 15 (n = 3). The final sample sizes per treatment and species are listed in @fig-results.

```{r, models}
#| label: models
# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
source(here("R", "func.R"))
# Model formula: FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id) + (1|clutch)
## A) L. delicata
deli <- fit_m(clean_df, "deli", "complete",  refit = FALSE)
write.csv(deli, file= "./output/Checking/deli.csv")
## B) L. guichenoti
guich <- fit_m(clean_df, "guich", "complete", refit = FALSE)
write.csv(guich, file= "./output/Checking/guich.csv")
```

```{r, results}
#| label: results
# Rename some of the posteriors and make means and sd for the posteriors.
## 1) L. delicata
deli_CORTCold <- deli$b_trial_reversal
deli_ControlCold <- (deli$'b_trial_reversal:cortControl' + deli$b_trial_reversal)
deli_CORTHot <- (deli$'b_trial_reversal:tempHot' + deli$b_trial_reversal)
deli_ControlHot <- (deli$'b_trial_reversal:cortControl:tempHot' + deli$b_trial_reversal + deli$'b_trial_reversal:cortControl' + deli$'b_trial_reversal:tempHot')
## 2) L. guichenoti
guich_CORTCold <- guich$b_trial_reversal
guich_ControlCold <- (guich$'b_trial_reversal:cortControl' + guich$b_trial_reversal)
guich_CORTHot <- (guich$'b_trial_reversal:tempHot' + guich$b_trial_reversal)
guich_ControlHot <- (guich$'b_trial_reversal:cortControl:tempHot' + guich$b_trial_reversal + guich$'b_trial_reversal:cortControl' + guich$'b_trial_reversal:tempHot')
```

Both species learnt across all treatments (@fig-results B, D), but we did not find any difference between treatments or species in the rate of learning (mean slopes per treatment for both species provided in Table 2 in Supplementary Material). For _L. delicata_, hormone treatment (slopes contrast: Control - CORT = `r format_dec(mean(c(deli_ControlCold, deli_ControlHot)) - mean(c(deli_CORTCold, deli_CORTHot)), 3)`, p~mcmc~ `r format_p(pmcmc(c(deli_ControlCold, deli_ControlHot) - c(deli_CORTCold, deli_CORTHot)), 3)`), incubation temperature (Hot - Cold = `r format_dec(mean(c(deli_CORTHot, deli_ControlHot)) - mean(c(deli_CORTCold, deli_ControlCold)),3)`, p~mcmc~ `r format_p(pmcmc(c(deli_CORTHot, deli_ControlHot) - c(deli_CORTCold, deli_ControlCold)),3)`), or the interaction ([(Control-Hot - CORT-Hot) - (Control-Cold - CORT-Cold)] = `r format_dec((mean(deli_ControlHot) - mean(deli_CORTHot)) - (mean(deli_ControlCold) - mean(deli_CORTCold)), 3)`, p~mcmc~ `r format_p(pmcmc((deli_ControlHot - deli_CORTHot) - (deli_ControlCold - deli_CORTCold)), 3)`) did not affect learning (see @fig-results A, B). We similarly found no significant effects of  hormone (Control - CORT = `r format_dec(mean(c(guich_ControlCold, guich_ControlHot)) - mean(c(guich_CORTCold, guich_CORTHot)), 3)`, p~mcmc~ `r format_p(pmcmc(c(guich_ControlCold, guich_ControlHot) - c(guich_CORTCold, guich_CORTHot)), 3)`), incubation temperature (Hot - Cold = `r format_dec(mean(c(guich_CORTHot, guich_ControlHot)) - mean(c(guich_CORTCold, guich_ControlCold)),3)`, p~mcmc~ `r format_p(pmcmc(c(guich_CORTHot, guich_ControlHot) - c(guich_CORTCold, guich_ControlCold)),3)`), or the interaction ([(Control-Hot - CORT-Hot) - (Control-Cold - CORT-Cold)] = `r format_dec((mean(guich_ControlHot) - mean(guich_CORTHot)) - (mean(guich_ControlCold) - mean(guich_CORTCold)), 3)`, p~mcmc~ `r format_p(pmcmc((guich_ControlHot - guich_CORTHot) - (guich_ControlCold - guich_CORTCold)), 3)`) on learning in _L. guichenoti_ (see @fig-results C, D).

```{r, fig-results}
#| label: fig-results
#| fig.cap: "Results for L. delicata (A,B) and L. guichenoti (C, D). Panels A and C show the predicted probability of choosing the correct feeder first over trials. The lines represent the mean predicted probability of choosing the correct feeder first on each trial, and the shaded areas indicate the standard deviation of the mean; both were obtained by using the slope and intercept estimates from the posterior distributions. The different colours indicate the different treatments. Panels B and D show the distribution of the estimates of slopes per each treatment. The x-axis represents the slope estimate, and in the y-axis are the density of the estimates. The different colours indicate the different treatments. Points and bars represent the mean and standard deviation of the mean of the estimates, respectively."
source(here("R", "func.R"))
# Create the dfs for the probability plot A and C) (df_prob)
# L. delicata
df_deliprob <- df_plotAC("deli")
# L. guichenoti
df_guichprob <- df_plotAC("guich")
#
# Create the df for the violin plots in plot B) (df_violin)
# L. delicata
df_deliV <- df_plotBD1("deli")
# L. guichenoti
df_guichV <- df_plotBD1("guich")
#
# Create the df for the points and bars in plot B) (df_points)
# L. delicata
df_deliP <- df_plotBD2("deli")
# L. guichenoti
df_guichP <- df_plotBD2("guich")
#
#
# Make plots deli and guich
source(here("R", "func.R"))
fig_deli <- plotting(sp = "deli", df_prob = df_deliprob, df_violin = df_deliV, df_points = df_deliP)
fig_guich <- plotting(sp = "guich", df_prob = df_guichprob, df_violin = df_guichV, df_points = df_guichP)
# Combine plots of both species into the final figure
fig_results <- plot_grid(fig_deli, fig_guich, ncol = 1)
ggsave("./output/figures/fig_results.png", plot=fig_results, width = 25, height = 18, units = "cm", dpi = 600)
knitr::include_graphics("./output/figures/fig_results.png")
```

Finally, when all individuals from all the treatments were pooled together and species were compared, there were not significant differences in the estimated slopes (mean of learning slope in _L. delicata_ = `r format_dec(mean(c(deli_CORTCold, deli_ControlCold, deli_CORTHot, deli_ControlHot)), 3)`, 95% CI = [`r format_dec(quantile(c(deli_CORTCold, deli_ControlCold, deli_CORTHot, deli_ControlHot), 0.025), 3)`, `r format_dec(quantile(c(deli_CORTCold, deli_ControlCold, deli_CORTHot, deli_ControlHot), 0.975), 3)`]; mean learning of slope in _L. guichenoti_ = `r format_dec(mean(c(guich_CORTCold, guich_ControlCold, guich_CORTHot, guich_ControlHot)), 3)`, 95% CI = [`r format_dec(quantile(c(guich_CORTCold, guich_ControlCold, guich_CORTHot, guich_ControlHot), 0.025), 3)`, `r format_dec(quantile(c(guich_CORTCold, guich_ControlCold, guich_CORTHot, guich_ControlHot), 0.975), 3)`]; _L. delicata_ - _L. guichenoti_ = `r format_dec(mean(c(deli_CORTCold, deli_ControlCold, deli_CORTHot, deli_ControlHot)) - mean(c(guich_CORTCold, guich_ControlCold, guich_CORTHot, guich_ControlHot)), 3)`, p~mcmc~ `r format_p(pmcmc(c(deli_CORTCold, deli_ControlCold, deli_CORTHot, deli_ControlHot) - c(guich_CORTCold, guich_ControlCold, guich_CORTHot, guich_ControlHot)), 3)`). Figures for both species with the raw data were included in the Supplementary Material.

## Discussion

This study aimed to determine whether prenatal temperature and GCs impact behavioural flexibility in two closely related species, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). We expected the early environment to influence behavioural flexibility in both species. We also predicted that _L. delicata_ would be more flexible than _L. guichenoti_ and less affected by prenatal conditions. Nevertheless, we did not find significant differences between treatments or species when we compared their performance in a reversal task. Our results show that, on average, all the lizards from the four treatments and the two species learnt the reversal task - indicated by estimated slopes that are significantly different from zero - and that they learn, on average, at the same rate. Furthermore, these results were consistent when only those individuals who performed better in the previous associative task were included (see Supplementary material).

Behavioural flexbility appears to be robust to early exposure to maternal GCs in both _L. delicata_ and _L. guichenoti_. We predicted that prenatal CORT treatment would decrease reversal learning abilities. As such, our results contrast with previous studies that show high GCs levels impaired performance in different types of learning tests [@lemaire_prenatal_2000; @zhu_prenatal_2004;  @farrell_developmental_2016]. Our results are, however, consistent with other experiments where GCs did not impact learning abilities [@szuran_water_1994; @bebus_associative_2016]. One possible reason for why GCs did not affect learning in our study may be because GCs can have hormetic effects on brain structure and learning. For example, in Sprague–Dawley rats, exposure to high doses of CORT, but not low doses, reduced the density of immature neurons and cell proliferation in the dentate gyrus [@DOSEneurobrummelte2010chronic]. It is possible such dose-dependent effects exist and our concentrations failed to elicit changes in the brain that would result in impaired behavioural flexibility. We find this possibility unlikely since, in a previous experiment, we observed that a similar dose affected a multitude of other traits including growth and baseline CORT levels in _L. delicata_ [@crino2024eggs]. Nonetheless, we cannot dismiss this possibility because dose effects can differ between tissues [@lemaire_prenatal_2000]. Conversely, the effects of GCs early in development may vary depending on the cognitive task, as evidenced by studies showing variations in the impact of prenatal GCs on different brain regions [@lemaire_prenatal_2000]. Prenatal GCs can have programmatic effects on HPA axis sensitivity, potentially impacting the lizards’ sensitivity to stressors [@crino_corticosterone_2014]. This altered stress response may influence learning dynamics through hormetic effects of CORT [@du_dynamic_2009], particularly in stressful contexts of which our experiment was not [@taborsky2021towards]. These hypotheses are not mutually exclusive and could be tested in future experiments using multiple doses of prenatal CORT treatment, assessing learning using multiple cognitive tasks, and testing the effect of stressors during the learning process.

More striking is the absence of an effect of incubation temperature on behavioural flexibility because most studies have reported better performance in different associative learning and motor tasks when incubated at higher temperatures [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. Studies that have reported cold-incubated lizards to learning faster [@dayananda_incubation_2017; @abayarathna_effects_2020] used incubation temperatures far beyond the natural range in their hot treatments and it is unclear how such conditions relate to wild environments. Our results suggest that behavioural flexibility is robust to temperature in both species of skinks - at least in this study. It could be argued that the differences in our incubation temperatures were not extreme enough to elicit significant differences in learning performance. However, the mean difference between the temperatures used in our experiment  were similar to those in previous studies [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014], and are aligned with natural variation at the extremes of incubation conditions in the wild [@qualls2000post; @cheetham2011embryonic]. In lizards, impacts on learning from early thermal environments are associated with changes in neural density [@amiel_effects_2017]. However, several studies have shown that the effects of temperature on brain structure and physiology are region dependent [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017]. It is possible that the temperatures used in our experiment did not affect the brain regions involved in coding reversal learning. Future studies should explore the effects of the thermal environment on brain structure and function in these species to understand the mechanisms underlying the effects of temperature on learning. Furthermore, testing learning across different cognitive tasks could provide a more comprehensive understanding of the influence of temperature on cognition. In our analyses, sex was not included as a factor because lizads were not considered sexually mature, and because previous studies have found weak evidence for sex-dependent learning in many cognitive domains [@szabo2019sex]. Indeed, the impacts of temperature on learning was not sex-depedent for the Eastern three-lined skink (_Bassiana dupeyerri_) [@amiel_hotter_2012; @clark_colour_2014]. However, some studies find evidence for interactive effects of incubation temperature and sex on brain physiology in other species of reptiles [@coomber_independent_1997; @sakata_neural_2000]. In addition, _B. dupeyerri_ are known to sex-reverse - some individuals can have female chromosomes but male phenotypes - under cold incubation temperatures [@dissanayake2021effects]. Such effects were not easily accounted for in these experiments making it challenging to understand if sex or temperature were the major drivers of learning differences [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014]. Further research should focus on the possible interactive effects of sex and incubation temperature on cognitive performance. However, such studies will be challenging given the large sample sizes required to execute these studies rigorously.

Despite sharing life-history traits, only _L. delicata_ has proven successful in colonising various global regions [@baker1979rainbow; @chapple2013divergent; @chapple2015deliinvLHI], and we predicted this may relate to higher behavioural flexibility. Furthermore, we expected behavioural flexibility in this species to be less susceptible to the effect of early environment since this trait can be adaptive in several stages of invasion involving more than one generation. While it will be important to explore replicate invasion events (both successful and unsuccessful) such information is seldom available. Nonetheless, there is precedence in the literature to expect differences between these two species [@chapple_know_2011; @chapple2013biosecurity]. Regardless, while we did not find differences in reversal learning between _L. delicata_ and _L. guichenoti_, previous studies have shown that _L. delicata_ more readily explores novel environments than _L. guichenoti_, which can be advantageous during colonisation [@chapple_know_2011]. However, @bezzina2014does found no differences in learning between the two skinks in an associative learning task. Our results align with the latter study and suggest that behavioural flexibility does not differ between these species either. Our results along with others' [@bezzina2014does] suggest that other aspects of their biology are more relevant for _L. delicata_'s success as invaders than learning and behavioural flexibility. Differences in exploratory behaviour is likely to be the key factor explaining why _L. delicata_ and not _L. guichenoti_ is better at colonising new areas [@chapple_know_2011].

In conclusion, our results indicate that lizards performed the reversal learning task, a measure of behavioural flexibility, at similar rates, with no discernible differences between treatments or species. These findings contrast with our initial predictions and suggest that behavioural flexibility in both species is robust to insults (at least for temperature and GCs) during early stages of development. Furthermore, our results indicate that other aspects of the biology of _L. delicata_ may be more influential for their success as invaders than behavioural flexibility alone. Nevertheless, future studies should investigate the potential influence of prenatal CORT and temperature on cognition, considering other factors such as the type of cognitive task. It is also imperative to explore the effects of these treatments on brain function comprehensively. Additionally, further research is essential to determine which aspects of _L. delicata_’s biology contribute more significantly to their success as invaders than behavioral flexibility alone. Investigating these areas will provide a better understanding of the cognitive and biological mechanisms underlying the adaptability and invasive potential of both species.

## Data accessibility  
All data, data description, and R code are available in public repository [https://github.com/Pablo-Recio/CORT_Temp_Behavioural_flexibility](https://github.com/Pablo-Recio/CORT_Temp_Behavioural_flexibility).

## Declaration of AI use
We declare Chat GPT was used for questions related to coding and data analyses. All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.

## Authors' contributions
P.R.: conceptualization, methodology, data collection, data curation, formal analysis, writing—original draft, writing—review and editing; D.C.L.: data collection, writing—review and editing; O.C.: conceptualization, methodology, writing—review and editing; K.H.W.: conceptualization, writing—review and editing; C.F.: conceptualization, methodology, funding acquisition, writing—review and editing; B.M: data collection, writing—review and editing; A.Y.P.: data collection, writing—review and editing; D.N.: conceptualization, methodology, funding acquisition, project administration, resources, supervision, writing—review and editing.  
All authors gave final approval for publication and agreed to be held accountable for the work performed therein.

## Conflict of interest declaration  
We declare we have no competing interests.

## Funding
This work was supported by a National Australian University fellowship (P.R.), the Australian Research Council (grant no. DP210101152) to D.N. and C.F., and the ACT Herpetological Association grant to P.R.

## Acknowledgements
We thank the help and assistance of our lab technicians Benjamin Durant and Michelle Stephens for taking care of the lizards. We are also grateful to ACTHA for the grant for the 3D printed feeders, and we also thank ANU MakerSpace, where we designed and built the prototypes of the 3D printed feeders. Finally, we wish to acknowledge the anonymous reviewers for their valuable feedback on the manuscript.

## References
<div id="refs"></div>

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

# Suplementary Material

Table 1. Estimates of Reversal learning slope for all the different treatments per each task, specie, and group. Mean shows the arithmetic means of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All pmcmc tested the hypothesis that the mean equals zero. In bold, those values that are significant (pmcmc <0.05).
```{r, tbl-data, results='asis'}
#| label: tbl-data
#| tbl-cap: "Estimates of Reversal learning slope for all the different treatments per each task, species, and group. Mean shows the arithmetic means of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All pmcmc tested the hypothesis that the mean equals zero. In bold, those values that are significant (pmcmc <0.05)"
source(here("R", "func.R"))
#
############################## CREATING BIG DF FOR TABLE ##############################
# Building the vectors for titles of rows and columns
specie <- c("L. delicata", "L. guichenoti")
treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
values <- c("Mean", "95% CI", "pmcmc")
# Building the vectors for estimated means, co.intervals(95%), and pmcmc for the slopes obtained from posteriors, assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
#First get estimates for both tasks
estimates <- list(
  deli_CORTCold, deli_ControlCold, deli_CORTHot, deli_ControlHot, 
  guich_CORTCold, guich_ControlCold, guich_CORTHot, guich_ControlHot
)
#
# Then get the mean, co.intervals(95%), and pmcmc
mean <- format_dec(sapply(estimates, mean), 3)
interval_025 <- format_dec(sapply(estimates, function(x) quantile(x,0.025)), 3)
interval_975 <- format_dec(sapply(estimates, function(x) quantile(x,0.975)), 3)
intervals <- paste(interval_025, interval_975, sep = " , ")
Pmcmc <- format_p(sapply(estimates, pmcmc), 4)
#
# Building the df
table_df <- data.frame(
  Specie = rep(specie, each = length(treatments)),
  Treatment = rep(rep(treatments, each = 1), times = length(specie)),
  Mean = rep(mean, each = 1),
  CI = rep(intervals, each = 1),
  Pmcmc = rep(Pmcmc, each = 1)
)
#
write.csv(table_df, file= "./output/Checking/df.csv")
#
############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
# Make n_list into a df
n_df <- as.data.frame(do.call(rbind, n_list)) %>%
  rename("n" = V1) %>%
  rownames_to_column("model") %>%
  separate(model, into = c("Specie", "cort", "temp"), sep = "_") %>%
  unite("Treatment", c("cort", "temp"), sep = "-") %>%
  mutate(Specie = factor(Specie,
                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
        Treatment = factor(Treatment,
                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
# Estimate total sample size per species and number of observations (this will allow to include this data in the first column of the table)
n_deli <- n_list$delicata_CORT_Cold+n_list$delicata_Control_Cold+n_list$delicata_CORT_Hot+n_list$delicata_Control_Hot
ob_deli <- sum(clean_df$species == "delicata", na.rm = TRUE)
n_guich <- n_list$guichenoti_CORT_Cold+n_list$guichenoti_Control_Cold+n_list$guichenoti_CORT_Hot+n_list$guichenoti_Control_Hot
ob_guich <- sum(clean_df$species == "guichenoti", na.rm = TRUE)
# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
new_table_df <- merge(table_df, n_df) %>%
  rename('pmcmc' = 'Pmcmc', '95% CI' = 'CI') %>% #Change the names of the columns for the table
  select(Specie, Treatment, Mean, `95% CI`, `pmcmc`, n) %>% #To order the columns in the way I want for the table
  mutate(Specie = factor(Specie,
                  levels = c("L. delicata", "L. guichenoti")),
        Treatment = factor(Treatment, 
                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
  arrange(Specie, Treatment) %>% # To arrange the rows the way I want
  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
  mutate(Treatment = paste0(Treatment, ")"))
write.csv(new_table_df, file= "./output/Checking/new_table_data.csv")
#
############################## MAKING THE TABLE ##############################
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 fint.size = 10)
# Split the table_data df by task
real_table <- flextable(new_table_df) %>%
    bold(~ `pmcmc` < 0.05, ~ `pmcmc` + Mean + `95% CI`) %>%
    set_table_properties(width = 1) %>%
    align(align="center", part="all") %>% 
    set_header_labels(Mean = "Mean",
                      `95% CI` = "95% CI",
                      `pmcmc` = "pmcmc") %>%
    italic(i= c(1, 5), j = 1, italic = TRUE, part = "body") %>% # To have names of species in italics
    flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
    flextable::compose(i = 2, j = 1, value = as_paragraph(sprintf("N = %d", n_deli)), part = "body") %>% # To write the delicata sample size
    flextable::compose(i = 6, j = 1, value = as_paragraph(sprintf("N = %d", n_guich)), part = "body") %>% # To write the guichenoti sample size
    flextable::compose(i = 3, j = 1, value = as_paragraph(sprintf("Obs = %d", ob_deli)), part = "body") %>% # To write the delicata total observations
    flextable::compose(i = 7, j = 1, value = as_paragraph(sprintf("Obs = %d", ob_guich)), part = "body") %>% # To write the guichenoti total observations
    hline(i = 4, j = c(1:5), part = "body") %>% # To make some horizontal lines
    autofit() 
real_table
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r, cleandata2}
#| label: cleandata2
# Cleaning the data to select only those individuals that passed a learning criterion of 80% correct choices in the last 10 trials of the associative task.
data <- read.csv(here("data/Learning.csv"))
percentage_per_lizard <- data %>%
  filter(trial_associative > 25 & trial_associative <= 35) %>%
  group_by(lizard_id) %>%
  summarise(percentage = mean(FC_associative == 1))
# Filter lizards where the percentage is at least 80%
filtered_lizards <- percentage_per_lizard %>%
  filter(percentage >= 0.8)
# Subset the original dataset based on the filtered lizard IDs
clean_df_2 <- data %>%
  filter(lizard_id %in% filtered_lizards$lizard_id) %>%
  filter(sum(is.na(FC_reversal)) <= 15) %>%
    mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>%
    mutate(temp = factor(temp,
      levels = c("23", "28"),
      labels = c("23" = "Cold", "28" = "Hot"))) %>%
    mutate(cort = factor(cort,
      levels = c("B", "A"),
      labels = c("B" = "CORT", "A" = "Control"))) %>%
    mutate(trial_reversal=as.numeric(trial_reversal)) %>%
  data.frame()
```

```{r, sampleSize2}
#| label: sampleSize2
# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
source(here("R", "func.R"))
#
specie <- c("delicata", "guichenoti")
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
#
n_list_2 <- list()
for(i in 1:length(specie)){
  for(k in 1:length(hormone)){
    for(l in 1:length(temperature)){
      n_list_2[[paste(specie[i], hormone[k], temperature[l], sep = "_")]] <- sample(clean_df_2, specie[i], hormone[k], temperature[l])
      }
    }
  }
```

```{r, models2}
#| label: models2
# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
source(here("R", "func.R"))
# Model formula: FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id) + (1|clutch)
## A) L. delicata
deli_2 <- fit_m(clean_df_2, "deli", "suppl", refit = FALSE)
write.csv(deli_2, file= "./output/Checking/deli.csv")
## B) L. guichenoti
guich_2 <- fit_m(clean_df_2, "guich", "suppl", refit = FALSE)
write.csv(guich_2, file= "./output/Checking/guich.csv")
```

```{r, resultsRev2}
# Rename some of the posteriors and make new estimates for the learning rate for the Reversal task doing the same thing we did in the chunk above.
## 1) L. delicata
deli_CORTCold_2 <- deli_2$b_trial_reversal
deli_ControlCold_2 <- (deli_2$'b_trial_reversal:cortControl' + deli_2$b_trial_reversal)
deli_CORTHot_2 <- (deli_2$'b_trial_reversal:tempHot' + deli_2$b_trial_reversal)
deli_ControlHot_2 <- (deli_2$'b_trial_reversal:cortControl:tempHot' + deli_2$b_trial_reversal + deli_2$'b_trial_reversal:cortControl' + deli_2$'b_trial_reversal:tempHot')
## 2) L. guichenoti
guich_CORTCold_2 <- guich_2$b_trial_reversal
guich_ControlCold_2 <- (guich_2$'b_trial_reversal:cortControl' + guich_2$b_trial_reversal)
guich_CORTHot_2 <- (guich_2$'b_trial_reversal:tempHot' + guich_2$b_trial_reversal)
guich_ControlHot_2 <- (guich_2$'b_trial_reversal:cortControl:tempHot' + guich_2$b_trial_reversal + guich_2$'b_trial_reversal:cortControl' + guich_2$'b_trial_reversal:tempHot')
```

Table 2. Estimates of Reversal learning slope for all the different treatments per each task, species, and group. Here we included only those individuals who made the right choice in 8 out of the last 10 trials in the previous associative task. Mean shows the arithmetic means of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All pmcmc tested the hypothesis that the mean equals zero. In bold, those values that are significant (pmcmc <0.05)

```{r, tbl-data_2}
#| label: tbl-data_2
#| tbl-cap: "Estimates of Reversal learning slope for all the different treatments per each task, species, and group. Here we included only those individuals who made the right choice in 8 out of the last 10 trials in the previous associative task. Mean shows the arithmetic means of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All pmcmc tested the hypothesis that the mean equals zero. In bold, those values that are significant (pmcmc <0.05)"
source(here("R", "func.R"))
#
############################## CREATING BIG DF FOR TABLE ##############################
# Building the vectors for titles of rows and columns
specie <- c("L. delicata", "L. guichenoti")
treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
values <- c("Mean", "95% CI", "pmcmc")
# Building the vectors for estimated means, co.intervals(95%), and pmcmcm for the slopes obtained from posteriors assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
#First get estimates for both tasks
estimates_2 <- list(
  deli_CORTCold_2, deli_ControlCold_2, deli_CORTHot_2, deli_ControlHot_2, 
  guich_CORTCold_2, guich_ControlCold_2, guich_CORTHot_2, guich_ControlHot_2
)
#
# Then get the mean, co.intervals(95%), and pmcmc
mean_2 <- format_dec(sapply(estimates_2, mean), 3)
interval_025_2 <- format_dec(sapply(estimates_2, function(x) quantile(x,0.025)), 3)
interval_975_2 <- format_dec(sapply(estimates_2, function(x) quantile(x,0.975)), 3)
intervals_2 <- paste(interval_025_2, interval_975_2, sep = " , ")
Pmcmc_2 <- format_p(sapply(estimates_2, pmcmc), 4)
#
# Building the df
table_df_2 <- data.frame(
  Specie = rep(specie, each = length(treatments)),
  Treatment = rep(rep(treatments, each = 1), times = length(specie)),
  Mean = rep(mean_2, each = 1),
  CI = rep(intervals_2, each = 1),
  Pmcmc = rep(Pmcmc_2, each = 1)
)
#
write.csv(table_df, file= "./output/Checking/df.csv")
#
############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
# Make n_list into a df
n_df_2 <- as.data.frame(do.call(rbind, n_list_2)) %>%
  rename("n" = V1) %>%
  rownames_to_column("model") %>%
  separate(model, into = c("Specie", "cort", "temp"), sep = "_") %>%
  unite("Treatment", c("cort", "temp"), sep = "-") %>%
  mutate(Specie = factor(Specie,
                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
        Treatment = factor(Treatment,
                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
# Estimate total sample size per species and number of observations (this will allow to include this data in the first column of the table)
n_deli_2 <- n_list_2$delicata_CORT_Cold+n_list_2$delicata_Control_Cold+n_list_2$delicata_CORT_Hot+n_list_2$delicata_Control_Hot
ob_deli_2 <- sum(clean_df_2$species == "delicata", na.rm = TRUE)
n_guich_2 <- n_list_2$guichenoti_CORT_Cold+n_list_2$guichenoti_Control_Cold+n_list_2$guichenoti_CORT_Hot+n_list_2$guichenoti_Control_Hot
ob_guich_2 <- sum(clean_df_2$species == "guichenoti", na.rm = TRUE)
# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
new_table_df_2 <- merge(table_df_2, n_df_2) %>%
  rename('pmcmc' = 'Pmcmc', '95% CI' = 'CI') %>% #Change the names of the columns for the table
  select(Specie, Treatment, Mean, `95% CI`, `pmcmc`, n) %>% #To order the columns in the way I want for the table
  mutate(Specie = factor(Specie,
                  levels = c("L. delicata", "L. guichenoti")),
        Treatment = factor(Treatment, 
                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
  arrange(Specie, Treatment) %>% # To arrange the rows the way I want
  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
  mutate(Treatment = paste0(Treatment, ")"))
write.csv(new_table_df, file= "./output/Checking/new_table_data.csv")
#
############################## MAKING THE TABLE ##############################
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 fint.size = 10)
# Split the table_data df by task
real_table_2 <- flextable(new_table_df_2) %>%
    bold(~ `pmcmc` < 0.05, ~ `pmcmc` + Mean + `95% CI`) %>%
    set_table_properties(width = 1) %>%
    align(align="center", part="all") %>% 
    set_header_labels(Mean = "Mean",
                      `95% CI` = "95% CI",
                      `pmcmc` = "pmcmc") %>%
    italic(i = c(1,5), j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
    flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
    flextable::compose(i = 2, j = 1, value = as_paragraph(sprintf("N = %d", n_deli_2)), part = "body") %>% # To write the delicata sample size
    flextable::compose(i = 6, j = 1, value = as_paragraph(sprintf("N = %d", n_guich_2)), part = "body") %>% # To write the guichenoti sample size
    flextable::compose(i = 3, j = 1, value = as_paragraph(sprintf("Obs = %d", ob_deli_2)), part = "body") %>% # To write the delicata total observations
    flextable::compose(i = 7, j = 1, value = as_paragraph(sprintf("Obs = %d", ob_guich_2)), part = "body") %>% # To write the guichenoti total observations
    hline(i = 4, j = c(1:5), part = "body") %>% # To make some horizontal lines
    autofit() 
real_table_2
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Figures with raw data

```{r, fig_rawdata}
#| label: fig_rawdata
#| fig.cap: "Raw data for the reversal tasks for both species. Lines represent the mean proportion of correct choices per trial for each treatment."
#
# Load the data
data_raw <- clean_df %>%
  mutate(trt = factor(trt, 
          levels = c("A_23", "B_23", "A_28", "B_28"), 
          labels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")),
        species = factor(species,
          levels = c("delicata", "guichenoti"),
          labels = c("italic('L. delicata')", "italic('L. guichenoti')"))) %>%
data.frame()
#
# Plot the raw data for the reversal tasks for both species
plot_raw <- ggplot(data_raw, aes(x = trial_reversal, y = FC_reversal, color = trt, fill = trt)) +
  geom_smooth(method = "loess", linewidth = 1, alpha = 0.1) +
  scale_color_manual(values = c("darkblue", "#68bde1", "#b50101", "#fa927d")) +
  scale_fill_manual(values = c("darkblue", "#68bde1", "#b50101", "#fa927d")) +
  facet_wrap(~species, nrow = 1, labeller = label_parsed) +
  labs(x = "Trial", y = "Proportion of correct choices", color = "Treatments", fill = "Treatments") +
  theme_classic()
#
# Save the plot
ggsave("./output/figures/plot_raw.png", plot=plot_raw, width = 25, height = 18, units = "cm", dpi = 600)
knitr::include_graphics("./output/figures/plot_raw.png")
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table 3. Number of individuals per treatment and species that reached a learning criterium of 5 consecutive correct choices (n lizards), and the average of trials taken (Trial) together with the standard deviation (sd).  
```{r, tbl-standard_crit, results='asis', out.widht = "26cm"}
#| label: standard_crit
#| tbl-cap: "Number of individuals per treatment and species that reached a learning criterium of 5 consecutive correct choices (n lizards), and the average of trials taken (Trial) together with the standard deviation (sd)."
#
source(here("R", "func.R"))
# Load the data
scrit_df <- clean_df %>%
  group_by(lizard_id,species, trt) %>%
  mutate(
    run_id = cumsum(lag(FC_reversal, default = 0) == 0 & FC_reversal == 1),
    run_length = ave(FC_reversal , run_id, FUN = cumsum) * FC_reversal 
  ) %>%
  filter(run_length >= 5) %>%
  group_by(lizard_id, species, trt) %>%
  slice_min(trial_reversal) %>%  # Get the first trial when the criterion was met
  summarize(
    species = first(species),
    trt = first(trt),
    trial_crit = min(trial_reversal),
    .groups = "drop") %>%
  mutate(trt = factor(trt, 
          levels = c("A_23", "B_23", "A_28", "B_28"), 
          labels = c("Control-Cold", "CORT-Cold", "Control-Hot", "CORT-Hot")),
        species = factor(species,
          levels = c("delicata", "guichenoti"),
          labels = c("L. delicata", "L. guichenoti")))%>%
  group_by(species, trt) %>%
  summarize(
    n = n(),
    mean_trial = format_dec(mean(trial_crit),2),
    des_trial = format_dec(sd(trial_crit),2),
    .groups = "drop") %>%
data.frame()
#
# Create the table
crit_tbl <- flextable(scrit_df) %>%
  set_table_properties(width = 1) %>%
  align(align = "center", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  set_header_labels(species = "Specie",
                    trt = "Treatment",
                    n = "n lizards",
                    mean_trial = "Mean",
                    des_trial = "sd") %>%
    italic(i = c(1,5), j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
    flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
    hline(i = 4, j = c(1:5), part = "body") %>% # To make some horizontal lines
    autofit()
crit_tbl
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Checking the effect of age on the learning rate in the associative task.

Model formula: 
Choice ~ age.start + trial_associative * cort * temp + (1 + trial_associative|lizard_id) + (1|clutch)

```{r, modelsage}
#| label: modelsage
# Fitting the models for L. delicata and L. guichenoti including age as a covariate
#
data_age <- clean_df %>%
  mutate(age.start = as.numeric(age.start)-mean(as.numeric(age.start)))
refit2 = FALSE
formula <- FC_reversal ~ age.start + trial_reversal*cort*temp + (1 + trial_reversal|lizard_id) + (1|clutch)
species <- c("delicata", "guichenoti")
#
models <- list() # Create an empty list to store the models
for(s in species){
  # Subset the data for the species
  data_species <- data_age %>%
    filter(species == s)
  # Define model file name
  model_file <- paste0("output/models/model_age_", s, ".rds")
# Fit the model only if it has not been fit yet (if refit2 = TRUE)
  if(refit2){
    # Fit the model
    model <- brm(formula,
                data = data_species,
                family = bernoulli(link = "logit"),
                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99))
    # Write the model to a file
    saveRDS(model, file = model_file)
  } else {
      # Read the model from a file
      model <- readRDS(file = model_file)
  }
  # Save the model to the list
  models[[s]] <- model
}
model_age_deli <- models[["delicata"]]
model_age_guich <- models[["guichenoti"]]
```

Table 4. Results for _L. delicata_ when the model included the age:
```{r, tbl-agedeli, results='asis', out.widht = "26cm"}
#| label: tbl-agedeli
#| tbl-cap: "Summary model_age_deli"
deli_age_res <- summary(model_age_deli)
# Extract fixed effects names and their corresponding estimates, CIs, etc.
fixed_effects <- deli_age_res$fixed %>%
  mutate_all(~ format_dec(., 2))
Predictors <- rownames(fixed_effects)
deli_age_tbl_df <- cbind(Predictors, fixed_effects)
# Convert the fixed effects data frame to a flextable
flextable(deli_age_tbl_df) %>%
  set_table_properties(., width = 1) %>%
  fontsize(size = 10, part = "all")
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Table 5. Results for _L. guichenoti_ when the model included the age:
```{r, tbl-ageguich, results='asis'}
#| label: tbl-ageguich
#| tbl-cap: "Summary model_age_guich"
guich_age_res <- summary(model_age_guich)
# Extract fixed effects names and their corresponding estimates, CIs, etc.
fixed_effects <- guich_age_res$fixed %>%
  mutate_all(~ format_dec(., 2))
Predictors <- rownames(fixed_effects)
guich_age_tbl_df <- cbind(Predictors, fixed_effects)
# Convert the fixed effects data frame to a flextable
flextable(guich_age_tbl_df)%>%
  set_table_properties(., width = 1) %>%
  fontsize(size = 10, part = "all")
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

```{r, fig-age} 
#| label: fig-age
#| fig.cap: "Distribution of the age of the lizards by treatment and species"
# Plotting the distribution of the age of the lizards by treatment and species
# Load the data
data_age <- clean_df
# Plot the distribution of the age of the lizards by treatment and species
img_deli <- readPNG("./Others/Deli.png")
plot_age_deli <- data_age %>%
  filter(species == "delicata") %>%
  mutate(trt = factor(trt, 
          levels = c("A_23", "B_23", "A_28", "B_28"), 
          labels = c("Control-Cold", "CORT-Cold", "Control-Hot", "CORT-Hot"))) %>%
  ggplot(aes(x = age.start, color = trt)) +
  geom_density(linewidth = 1.5) +
  scale_color_manual(values = c("CORT-Cold"="darkblue", "Control-Cold"="#68bde1", "CORT-Hot"="#b50101", "Control-Hot"="#fa927d")) +  
  labs(title = "L. delicata", x = "Age (days)", color = "Treatments") +
  theme_classic() +
  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"),
    plot.title = element_text(size = 14, family = "Times", face = "italic"),
    axis.text.y = element_blank(),   # Remove y-axis labels
    axis.title = element_text(size = 12, family = "Times"),
    axis.text = element_text(size = 10, family = "Times"),
    legend.position = "right",
    legend.title = element_text(size = 12, family = "Times"),
    legend.text = element_text(size = 11, family = "Times")
    ) +
  annotation_custom(rasterGrob(img_deli), xmin = 0.73, xmax = 0.98, ymin = 0.73, ymax = 0.98)
#
img_guich <- readPNG("./Others/Guich.png")
plot_age_guich <- data_age %>%
  filter(species == "guichenoti") %>%
  mutate(trt = factor(trt, 
          levels = c("A_23", "B_23", "A_28", "B_28"), 
          labels = c("Control-Cold", "CORT-Cold", "Control-Hot", "CORT-Hot"))) %>%
  ggplot(aes(x = age.start, color = trt)) +
  geom_density(linewidth = 1.5) +
  scale_color_manual(values = c("CORT-Cold"="darkblue", "Control-Cold"="#68bde1", "CORT-Hot"="#b50101", "Control-Hot"="#fa927d")) +  
  labs(title = "L. guichenoti", x = "Age (days)", color = "Treatments") +
  theme_classic() +
  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"),
    plot.title = element_text(size = 14, family = "Times", face = "italic"),
    axis.text.y = element_blank(),   # Remove y-axis labels
    axis.title = element_text(size = 12, family = "Times"),
    axis.text = element_text(size = 10, family = "Times"),
    legend.position = "right",
    legend.title = element_text(size = 12, family = "Times"),
    legend.text = element_text(size = 11, family = "Times")
    ) +
  annotation_custom(rasterGrob(img_guich), xmin = 0.73, xmax = 0.98, ymin = 0.73, ymax = 0.98)
#
plot_age <- plot_grid(plot_age_deli, plot_age_guich, ncol = 1)
ggsave("./output/figures/plot_age.png", plot=plot_age, width = 25, height = 18, units = "cm", dpi = 600)
knitr::include_graphics("./output/figures/plot_age.png")
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Checking the main models plots

```{r, residuals}
# Chunk for calling all the models and making the residuals
## L. delicata
mod_deli <- readRDS(here("output/models/delicomplete.rds"))
resid_deli <- residuals(mod_deli)
## L. guichenoti
mod_guich <- readRDS(here("output/models/guichcomplete.rds"))
resid_guich <- residuals(mod_guich)
```

Model formula for the reversal task is:

Choice  ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)

Plots for the different models of the associative task:  
1.- _L. delicata_    
```{r , plotmod_drr, out.width="70%", fig.align="center"}
bayes_R2(mod_deli)
plot(mod_deli)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

2.- _L. guichenoti_  
```{r , plotmod_grr, out.width="70%", fig.align="center"}
bayes_R2(mod_guich)
plot(mod_guich)
```